---
title: "DATA624 Homework 1"
author: "Esteban Aramayo, Coffy Andrews-Guo, LeTicia Cancel, Joseph Connolly, Ian Costello"
date: "5/31/2022"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
---

```{r warning=FALSE, error=FALSE, message=FALSE}
library(ggfortify)
library(openxlsx)
library(fpp2)
library(fma)
library(gridExtra)
library(seasonal)
library(ggplot2)
library(patchwork)
```

# Week 1

## HA 2.1 

### Use the help function to explore what the series `gold`, `woolyrnq` and `gas` represent.

```{r}
#help function for each series
??gold
```

```{r}
??woolyrnq
```

```{r}
??gas
```

### a. Use `autoplot()` to plot each of these in separate plots.

```{r fig.height=15, fig.width=10}
#autoplots of each series
grid.arrange(autoplot(gold),autoplot(woolyrnq),autoplot(gas))
```

### b. What is the frequency of each series? Hint: apply the `frequency()` function.

- `gold` has a frequency of 1, meaning it is annual.
- `Woolyrng` has a frequency of 1, meaning it is quarterly. 
- `gas` has a frequency of 12, meaning it is monthly.

```{r}
#frequency of each series
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

### c. Use `which.max()` to spot the outlier in the `gold` series. Which observation was it?
 
The outlier for the `gold` series is observation number 770, with the value of 593.7.

```{r}
which.max(gold)
gold[which.max(gold)]
```

## HA 2.3

Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

### a. You can read the data into R with the following script:
```{r}
retaildata <- read.xlsx("https://otexts.com/fpp2/extrafiles/retail.xlsx",startRow = 2)
```

### b. Select one of the time series as follows (but replace the column name with your own chosen column):
```{r}
myts <- ts(retaildata[,"A3349873A"],
  frequency=12, start=c(1982,4))
myts

myts2 <- ts(retaildata[,"A3349791W"],
            frequency=12, start=c(1982,4))
myts2
```

#### Explore your chosen retail time series using the following functions:

`autoplot()`, `ggseasonplot()`, `ggsubseriesplot()`, `gglagplot()`, `ggAcf()`

#### Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

The first plot using autoplot() shows some seasonality for each year telling us that the price spikes at different points. There is also an upward trend showing that the price of gold continues to increase overtime. The length between the min and max for each of the seasonal spike grows longer each year telling us that the price difference between the minimum and maximum price spreads out more over time. 
```{r}
#time series qutoplot
autoplot(myts2)
```

The seasonal plot below using ggseasonlot() gives us more clarity on when the seasonality takes place. Each year has a spike in December with the max prices increasing each year. We can take a guess as to why December is the month that spikes since this is when holiday shopping takes place. 

```{r}
#time series seasonal plot
ggseasonplot(myts2)
```

The subseries plot confirms the seasonality shown in the previous plots with December being the seasonal month. This plot creates a better visualization for a couple of details better than the previous plots. The difference between the minimum and maximum over the years is clearer. The minimum has a minor increase in december but the maximum has a dramatic increase when compared to the other months. The mean also has a big increase in December but then levels out during the spring and summer months.

```{r}
#time series sub series plot
ggsubseriesplot(myts2)
```

The lag plot shows strong positive correlation for all months in lag 12.

```{r fig.width=15}
#time series lag plot
gglagplot(myts2)
```

The ACF plot supports the lag plot by showing a high autocorrelation coefficient for lag 12. There are also no negative correlations which was not as clear in the subseries plot. They are also all significantly different from zero. 

```{r}
#time series ACF
ggAcf(myts2)

```

## HA 6.2
The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

### a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?


```{r fig.width=10, fig.height=15}
#autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()
p1 <- autoplot(plastics)
p2 <- ggseasonplot(plastics)
p3 <- ggsubseriesplot(plastics)
p4 <- gglagplot(plastics)
p5 <- ggAcf(plastics)
grid.arrange(p1,p2,p3,p5,p4, ncol=1)
```

### b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r fig.height=5, warning=FALSE}
#multiplicative decomposition
decomp_plastics <- plastics %>%
  decompose(type = "multiplicative")

decomp_plastics %>% autoplot()
```

### c. Do the results support the graphical interpretation from part a?

The "Data" and "Seasonal" sections of the Decomposition shows similar results to the autoplot() chart. There is a steady seasonal trend that has a similar duration for each seasonal cycle. The "Trend" section of the Decomposition also supports the visuals from part A and shows a steady upward trend from years 1-5. 

### d. Compute and plot the seasonally adjusted data.

```{r warning=FALSE}
autoplot(plastics, series = "Data") +  
  autolayer(trendcycle(decomp_plastics), series = "Trend") +
  autolayer(seasadj(decomp_plastics), series = "Seasonally Adjusted")

#fit <- seas(x = plastics, x11="")

plastics %>%
  stl(s.window = "periodic", robust = TRUE) %>%
  autoplot()
```

### e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?
The outlier was applied to the first observation in the plastics data so the chart changed only in year 1. The outlier being in the beginning of the data had no impact on the rest of the plot.
```{r warning=FALSE}
#creating a copy of plastics and adding 500 to one observation
plastics2 <- plastics
plastics2[1] <- plastics2[1] + 500

decomp_plastics2 <- plastics2 %>%
  decompose(type = "multiplicative")

autoplot(plastics2, series = "Data") +  
  autolayer(trendcycle(decomp_plastics2), series = "Trend") +
  autolayer(seasadj(decomp_plastics2), series = "Seasonally Adjusted")
```

### f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?

The middle outlier has a greater impact on the Seasonally adjusted line than the end outlier when compared to the original plot. When the last observation is the outlier the Data and Seasonally Adjusted lines trend upward and all other areas on the chart are not impacted. When a middle observation is the outlier we see a different seasonally adusted line when compared to the original chart. The downward and upward spikes around the outlier are sharper on the middle outlier chart while the original chart has subtle changes with a fairly even upward trend. 
```{r fig.height=10, warning=FALSE}
plastics_mid <- plastics
plastics_end <- plastics
plastics_mid[30] <- plastics_mid[30] + 500
plastics_end[60] <- plastics_end[60] + 500

decomp_mid <- plastics_mid %>%
  decompose(type = "multiplicative")

p <- autoplot(plastics, series = "Data") +  
  autolayer(trendcycle(decomp_plastics), series = "Trend") +
  autolayer(seasadj(decomp_plastics), series = "Seasonally Adjusted") +
  ggtitle("Plastics Seasonally Adjusted - Original")

p1 <- autoplot(plastics_mid, series = "Data") +  
  autolayer(trendcycle(decomp_mid), series = "Trend") +
  autolayer(seasadj(decomp_mid), series = "Seasonally Adjusted") +
  ggtitle("Plastics Seasonally Adjusted - Mid Observation")

decomp_end <- plastics_end %>%
  decompose(type = "multiplicative")

p2 <- autoplot(plastics_end, series = "Data") +  
  autolayer(trendcycle(decomp_end), series = "Trend") +
  autolayer(seasadj(decomp_end), series = "Seasonally Adjusted") +
  ggtitle("Plastics Seasonally Adjusted - End Observation")

grid.arrange(p, p1,p2)
```

# Week 2

## KJ 3.1

The UC Irvine Machine Learning Repository contains a data set related to glass identification. The data consist of 214 glass samples labeled as one of seven class categories. There are nine predictors, including the refractive index and percentages of eight elements: Na, Mg, Al, Si, K, Ca, Ba, and Fe. The data can be accessed via:
```{r warning=FALSE, error=FALSE}
library(mlbench)
data(Glass)
str(Glass)
```

### a. Using visualizations, explore the predictor variables to understand their distributions as well as the relationships between predictors.
When we explore the Glass dataset using summary we see that there are no NA values for any of the predictors. There are also no negative values with the minimum value being 0.00 (zero) for some predictors. 
```{r}
summary(Glass)
```

```{r fig.width=10, fig.height=10}
ri_b <- ggplot(Glass, aes(x = Type, y = RI)) +
  geom_boxplot()

na_b <- ggplot(Glass, aes(x = Type, y = Na)) +
  geom_boxplot()

mg_b <- ggplot(Glass, aes(x = Type, y = Mg)) +
  geom_boxplot()

al_b <- ggplot(Glass, aes(x = Type, y = Al)) +
  geom_boxplot()

si_b <- ggplot(Glass, aes(x = Type, y = Si)) +
  geom_boxplot()

k_b <- ggplot(Glass, aes(x = Type, y = K)) +
  geom_boxplot()

ca_b <- ggplot(Glass, aes(x = Type, y = Ca)) +
  geom_boxplot()

ba_b <- ggplot(Glass, aes(x = Type, y = Ba)) +
  geom_boxplot()

fe_b <- ggplot(Glass, aes(x = Type, y = Fe)) +
  geom_boxplot()


ri_b+na_b+mg_b+al_b+si_b+k_b+ca_b+ba_b+fe_b+
  plot_layout(ncol=3)


```

### b. Do there appear to be any outliers in the data? Are any predictors skewed?

The boxplots help us see the outliers for each type for each predictor. 
```{r}

```

### c. Are there any relevant transformations of one or more predictors that might improve the classification model?
```{r}

```

## KJ 3.2

The soybean data can also be found at the UC Irvine Machine Learning Repository. Data were collected to predict disease in 683 soybeans. The 35 predictors are mostly categorical and include information on the environmental conditions (e.g., temperature, precipitation) and plant conditions (e.g., left spots, mold growth). The outcome labels consist of 19 distinct classes. The data can be loaded via:
```{r}
library(mlbench)
data(Soybean)
## See ?Soybean for details
```

### a. Investigate the frequency distributions for the categorical predictors. Are any of the distributions degenerate in the ways discussed earlier in this chapter?
### b. Roughly 18 % of the data are missing. Are there particular predictors that are more likely to be missing? Is the pattern of missing data related to the classes?
### c. Develop a strategy for handling missing data, either by eliminating predictors or imputation.

## KJ 635

## HA 7.1

### Consider the pigs series — the number of pigs slaughtered in Victoria each month.

```{r}
help(pigs)
pigs %>% autoplot()
```

### a. Use the `ses()` function in R to find the optimal values of \(\alpha\) and \(l_0\) and generate forecasts for the next four months.

Using the `summary()` function $\alpha = 0.2971$ and $l_0 = 77260.0561$.

```{r}
fc<- ses(pigs, h=4)

fc %>%
  autoplot()

summary(fc)
```

#### b. Compute a 95% prediction interval for the first forecast using \(\hat{y}\pm 1.96s\) where \(s\) is the standard deviation of the residuals. Compare your interval with the interval produced by R.

From chapter 3.5. Prediction intervals are calculated as $\hat{y}_T+h|T \pm c \hat{\sigma}_h$. The multiplier for 95% interval is 1.96 and the residuals are also equal to the RMSE. Using the SES model formula from part a, the $\hat{y} values are stored in a vector as are the residuals. Then, the values are subbed in for the formula. 

The computed intervals vs. the predicted intervals are quite close. The September 1995 low interval at 95% is 78611.97, compared to the computed 78679.97 only a difference of 68. The high values for the predicted interval at 95% is 119020.80, compared to 118952.84 again differing by 68.

```{r}
y_hat <- c(1.96, -1.96)
s <- sd(residuals(fc))

ses(pigs, h=4)$mean[1]+(y_hat*s)
```

## HA 7.2 

## HA 7.3


