---
title: "Data624 - Project 1"
author: "Esteban Aramayo, Coffy Andrews-Guo, LeTicia Cancel, Joseph Connolly, Ian Costello"
date: '2022-06-21'
output: 
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE,
                      message=FALSE,
                      collapse = FALSE,
                      comment = "#>" )
```

```{r echo=FALSE}
# install from CRAN
#install.packages("officedown")

# or GitHub
#remotes::install_github("davidgohel/officedown")
```

Required Libraries
```{r warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(fpp2)
library(caret)
library(RANN)
library(VIM)
library(ggpubr)
library(gridExtra)
library(forecast)
```

# Project Summary

De-identified data was provided to conduct a series of six forecasts of different variables of a provided data set. There are two major requirements of this project:

1. This written report
2. The forecasts and error rates

```{r}
df <- read_excel("data.xls")
head(df)
```



## Data Cleaning & Imputation

```{r}
# Factoring category to get a count of the elements within dataset
df$category <- as.factor(df$category)

summary(df)
```

- All groups are equal in length

- Columns 5-7 (Var03, Var04, Var07) all have same amount of missing values. Very close quartile and min/max values, comparable to column 3 (Var01). Column 4 (Var02) has values that are significantly larger 

#### Data Structure
```{r}
str(df)
```


## Data Exploration

```{r}
dim(df)
```

```{r}
summary(df)
```

- The provided data set contains 7 columns and 10,572 rows. The first column, "SeriesInd", is a datetime variable which can be converted to reflect a date. "category" is a classification of each row, of which there are 6 different ones, and the remaining 5 columns are variables to forecast and are the key ingredient for constructing timeseries.

##### Handling Missing Data

```{r}
paste0(sum(is.na(df))," values missing from original set")
```

Looking at the summary generated above, columns 3-7 each have a range of 842-866 missing values. The dilemma is to decide whether or not it is appropriate to perform an analysis via imputing missing values, or to simply delete them. According to the plot below, generated via "VIM::aggr()", 91.81% of the data is fulfilled. Var01, Var02, Var03, Var05, and Var07 are missing about 8% of data. At first impression, this seems like an insignificant amount of data that can be omitted from the set. Further investigation is needed to confirm this impression, and deletion will be deemed appropriate if and only the data is found to be missing completely at random (MCAR).

##### Impute or Delete?

An excerpt from the following paper, $\underline{The\ prevention\ and\ handling\ of\ the\ missing\ data}$, by Hyun Kang, argues when deletion is appropriate or not from the following quote: *"...if the assumption of MCAR (missing completely at random) is satisfied, a listwise deletion is known to produce unbiased estimates and conservative results. When the data do not fulfill the assumption of MCAR, listwise deletion may cause bias in the estimates of the parameters. If there is a large enough sample, where power is not an issue, and the assumption of MCAR is satisfied, the listwise deletion may be a reasonable strategy. However, when there is not a large sample, or the assumption of MCAR is not satisfied, the listwise deletion is not the optimal strategy"* [$^1$](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3668100/)

```{r fig.width=10}
# Plots of missing values

aggr_plot <- VIM::aggr(df, col = c("navyblue", "orange"), 
                  numbers = T, sortVars = T,
                  labels = names(df),
                  cex.axis = 0.7, gap = 3,
                  ylab = c("Frequency of Missing Data", "Pattern"))
```


By creating a shadow matrix to see a percentage (on a 0-1 scale) of missing values when all correlated among each other, this will help indicate whether or not the data is MCAR. [$^2$](https://stats.stackexchange.com/questions/172316/a-statistical-approach-to-determine-if-data-are-missing-at-random)

```{r}
# Shadow Matrix: correlation of missing values from the dataset

x <- as.data.frame(abs(is.na(df))) 

y <- x[which(sapply(x, sd) >0)] # Extracts which variables are missing/NA from the dataset

cor(y) # Tendency of NA when correlated among variables
```

Aside from considering values correlated with themselves, the following have no missing values when correlated with others:

  - Var03 has no missing values when correlated with Var05 and Var07
  
  - Var05 has no missing values when correlated with Var03 and Var07
  
  - Var07 has no missing values when correlated with Var03 and Var05
  
Taking these observations into consideration, it seems there appears to be bias in the data in the context of missing data and therefore not missing completely at random. Instead of a listwise deletion, an imputation will be performed

##### Data Imputation

- Imputation via "medianImpute" method within "preProcess()" function via the caret package will be utilized. 

```{r}
preProcess_NAdata_model <- preProcess(as.data.frame(df), method ="medianImpute")

df <- predict(preProcess_NAdata_model, newdata = df)

paste0(sum(is.na(df))," values missing after imputation")
```

- No data is missing after imputation
- No discrepencies in the data after imputation.

```{r}
summary(df)
```


##### Datetime conversion

- Datetime conversion is to be performed after imputation in order to ensure a successful imputation

```{r}
# Converting Var02 to Datetime
df$SeriesInd <- as.integer(df$SeriesInd)
df$SeriesInd <- as.POSIXct(df$SeriesInd, origin = "1970-01-01")


# Renaming SeriesInd to Date to clarify purpose

df <- df %>% rename("Datetime" = SeriesInd)
summary(df)
```

##### Final susbets 

```{r}
# For forecasting later on

s01 <- df %>% filter(category == "S01")
s02 <- df %>% filter(category == "S02")
s03 <- df %>% filter(category == "S03")
s04 <- df %>% filter(category == "S04")
s05 <- df %>% filter(category == "S05")
s06 <- df %>% filter(category == "S06")
```

## Data Exploration

The data set does not appear to have any distinguishing labels that would indicate anything about the source or purpose of the data set. Under normal circumstances, context about data and use case is important for forecasting. Context may help identify the kinds of methods and models that would be best suited to produce an accurate result - following the "no free lunch" principle. 

Boxplots: Var02 has the most outliers but this is also the column with the largest range in values

```{r warning=FALSE, fig.width=10, fig.height=8}
p1 <- ggplot(df, aes(category, Var01)) +
  geom_boxplot()
p2 <- ggplot(df, aes(category, Var02)) +
  geom_boxplot()
p3 <- ggplot(df, aes(category, Var03)) +
  geom_boxplot()
p4 <- ggplot(df, aes(category, Var05)) +
  geom_boxplot()
p5 <- ggplot(df, aes(category, Var07)) +
  geom_boxplot()
```


## Data Visualization - Predicators

A time series that contain a list of numbers (the measurements), along with some information about what times those numbers were recorded (the index). 


*Box plots*

- Var02 has the most outliers but this is also the column with the largest range in values
```{r warning=FALSE, message=FALSE, fig.width=6, fig.height=8}
p1 <- ggplot(df, aes(Var01, fill = category)) +
  geom_boxplot(outlier.color = "red", outlier.size = 3)
p2 <- ggplot(df, aes(Var02, fill = category)) +
  geom_boxplot(outlier.color = "red", outlier.size = 3)
p3 <- ggplot(df, aes(Var03, fill = category)) +
  geom_boxplot(outlier.color = "red", outlier.size = 3)
p4 <- ggplot(df, aes(Var05, fill = category)) +
  geom_boxplot(outlier.color = "red", outlier.size = 3)
p5 <- ggplot(df, aes(Var07, fill = category)) +
  geom_boxplot(outlier.color = "red", outlier.size = 3)
p6 <- ggplot(df, aes(Datetime, fill = category)) + 
  geom_boxplot(outlier.color = "red", outlier.size = 10)

(p1+p2)/(p3+p4)/(p5+p6)
```

All predictors are left skewed so transformations are needed for each column

```{r warning=FALSE, fig.width=10, fig.height=7}
p1 <- ggplot(df, aes(Var01, fill=category)) +
  geom_density(alpha = 0.5)
p2 <- ggplot(df, aes(Var02, fill=category)) +
  geom_density(alpha = 0.5)
p3 <- ggplot(df, aes(Var03, fill=category)) +
  geom_density(alpha = 0.5)
p4 <- ggplot(df, aes(Var05, fill=category)) +
  geom_density(alpha = 0.5)
p5 <- ggplot(df, aes(Var07, fill=category)) +
  geom_density(alpha = 0.5)

p1+p2+p3+p4+p5+
  plot_layout(ncol = 2)
```

## Data transformations

Var02 is the most skewed
```{r}
library(moments)
skewness(df$Var01)
skewness(df$Var02)
skewness(df$Var03)
skewness(df$Var05)
skewness(df$Var07)
```

Experimenting with 3 different transformations; log, square root, and cube root. The log transformation looks the most normalized

```{r error=FALSE}
log_var01 <- log10(df$Var01)
sqrt_var01 <- sqrt(df$Var01)
cube_var01 <- df$Var01^(1/3)

hist(df$Var01)
hist(log_var01)
hist(sqrt_var01)
hist(cube_var01)


```

Transform all predictors using log transformation

```{r}
df_transformed <- df
df_transformed$Var01 <- log10(df$Var01)
df_transformed$Var02 <- log10(df$Var02)
df_transformed$Var03 <- log10(df$Var03)
df_transformed$Var05 <- log10(df$Var05)
df_transformed$Var07 <- log10(df$Var07)
```

New plots using transformed dataframe. Var02 is the most normalize so we will try different transformations with the other columns

```{r warning=FALSE, fig.width=10, fig.height=7}
p1 <- ggplot(df_transformed, aes(Var01, color=category)) +
  geom_density()
p2 <- ggplot(df_transformed, aes(Var02, color=category)) +
  geom_density()
p3 <- ggplot(df_transformed, aes(Var03, color=category)) +
  geom_density()
p4 <- ggplot(df_transformed, aes(Var05, color=category)) +
  geom_density()
p5 <- ggplot(df_transformed, aes(Var07, color=category)) +
  geom_density()

p1+p2+p3+p4+p5+
  plot_layout(ncol = 2)
```

```{r}
s01 <- df_transformed %>% dplyr::filter(category == "S01")
s02 <- df_transformed %>% dplyr::filter(category == "S02")
s03 <- df_transformed %>% dplyr::filter(category == "S03")
s04 <- df_transformed %>% dplyr::filter(category == "S04")
s05 <- df_transformed %>% dplyr::filter(category == "S05")
s06 <- df_transformed %>% dplyr::filter(category == "S06")
```

Conversion using at.Date so each row is a date.
```{r}
df_test <- read_excel("data.xls")
head(df_test)
tail(df_test)

# Converting Var02 to Datetime
df_test$SeriesInd <- as.Date(df_test$SeriesInd, origin = "1899-12-30")


# Renaming SeriesInd to Date to clarify purpose
df_test <- df_test %>% rename("Date" = SeriesInd)
summary(df_test)

#new imputation
preProcess_NAdata_model <- preProcess(as.data.frame(df_test), method ="medianImpute")

df_test <- predict(preProcess_NAdata_model, newdata = df_test)

paste0(sum(is.na(df_test))," values missing after imputation")

#new subsets with data conversion
s01_2 <- df %>% filter(category == "S01")
s02_2 <- df %>% filter(category == "S02")
s03_2 <- df %>% filter(category == "S03")
s04_2 <- df %>% filter(category == "S04")
s05_2 <- df %>% filter(category == "S05")
s06_2 <- df %>% filter(category == "S06")
```


create time series for each subset using the dataframe with dates
```{r}
s01_ts <- ts(s01_2[,c("Var01","Var02")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s02_ts <- ts(s02_2[,c("Var02","Var03")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s03_ts <- ts(s03_2[,c("Var05","Var07")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s04_ts <- ts(s04_2[,c("Var01","Var02")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s05_ts <- ts(s05_2[,c("Var02","Var03")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s06_ts <- ts(s06_2[,c("Var05","Var07")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
```


Autoplots and seasonal plots for S01
```{r fig.width=15, fig.height=15}
(autoplot(s01_ts, facets = 2)) / 
  (ggseasonplot(s01_ts[,1], main = "Var01") + ggseasonplot(s01_ts[,2], main = "Var02")) / 
  (ggsubseriesplot(s01_ts[,1]) + ggsubseriesplot(s01_ts[,2])) / 
  (ggAcf(s01_ts[,2], main = "") + ggAcf(s01_ts[,2], main = ""))
```

Autoplots and seasonal plots for S02
```{r fig.width=15, fig.height=15}
(autoplot(s02_ts, facets = 2)) / 
  (ggseasonplot(s02_ts[,1], main = "Var02") + ggseasonplot(s02_ts[,2], main = "Var03")) / 
  (ggsubseriesplot(s02_ts[,1]) + ggsubseriesplot(s02_ts[,2])) / 
  (ggAcf(s02_ts[,2], main = "") + ggAcf(s02_ts[,2], main = ""))
```

Autoplots and seasonal plots for S03
```{r fig.width=15, fig.height=15}
(autoplot(s03_ts, facets = 2)) / 
  (ggseasonplot(s03_ts[,1], main = "Var05") + ggseasonplot(s03_ts[,2], main = "Var07")) / 
  (ggsubseriesplot(s03_ts[,1]) + ggsubseriesplot(s03_ts[,2])) / 
  (ggAcf(s03_ts[,2], main = "") + ggAcf(s03_ts[,2], main = ""))
```

Autoplots and seasonal plots for S04
```{r fig.width=15, fig.height=15}
(autoplot(s04_ts, facets = 2)) / 
  (ggseasonplot(s04_ts[,1], main = "Var01") + ggseasonplot(s04_ts[,2], main = "Var02")) / 
  (ggsubseriesplot(s04_ts[,1]) + ggsubseriesplot(s04_ts[,2])) / 
  (ggAcf(s04_ts[,2], main = "") + ggAcf(s04_ts[,2], main = ""))
```

Autoplots and seasonal plots for S05
```{r fig.width=15, fig.height=15}
(autoplot(s05_ts, facets = 2)) / 
  (ggseasonplot(s05_ts[,1], main = "Var02") + ggseasonplot(s05_ts[,2], main = "Var03")) / 
  (ggsubseriesplot(s05_ts[,1]) + ggsubseriesplot(s05_ts[,2])) / 
  (ggAcf(s05_ts[,2], main = "") + ggAcf(s05_ts[,2], main = ""))
```

Autoplots and seasonal plots for S06
```{r fig.width=15, fig.height=15}
(autoplot(s06_ts, facets = 2)) / 
  (ggseasonplot(s06_ts[,1], main = "Var05") + ggseasonplot(s06_ts[,2], main = "Var07")) / 
  (ggsubseriesplot(s06_ts[,1]) + ggsubseriesplot(s06_ts[,2])) / 
  (ggAcf(s06_ts[,2], main = "") + ggAcf(s06_ts[,2], main = ""))
```


```{r echo=TRUE, eval = FALSE, results='hide'}
p1 <- (gglagplot(s01_ts[,1]) + theme(legend.position = "none") + gglagplot(s01_ts[,2]) + theme(legend.position = "none") )
p2 <- gglagplot(s02_ts[,1]) + theme(legend.position = "none") + gglagplot(s02_ts[,2]) + theme(legend.position = "none")
p3 <- gglagplot(s03_ts[,1]) + theme(legend.position = "none") + gglagplot(s03_ts[,2])
p4 <- gglagplot(s04_ts[,1]) + theme(legend.position = "none") + gglagplot(s04_ts[,2]) + theme(legend.position = "none")
p5 <- gglagplot(s05_ts[,1]) + theme(legend.position = "none") + gglagplot(s05_ts[,2]) + theme(legend.position = "none")
p6 <- gglagplot(s06_ts[,1]) + theme(legend.position = "none") + gglagplot(s06_ts[,2]) 

```
```{r fig.width=20}
grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 2)
```

## Data Modeling and Forecasting

### Forecasting with Decomposition

#### Forecasting S01: Var01 & Var02 with decomposition

```{r fig.height=10, fig.width=15}
#STL using default values
fit_stl_1 <- stl(s01_ts[,1], s.window = "periodic")

#STL using default values
fit_stl_2 <- stl(s01_ts[,2], s.window = "periodic")

#forecast of seasonaly adjusted data
f1 <- fit_stl_1 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S01: Var01")

#forecast of seasonaly adjusted data
f2 <- fit_stl_2 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S01: Var02")

#forecast from SLT + Random walk
f3 <- fit_stl_1 %>% forecast(method="naive") %>%
  autoplot() + ylab("S01: Var01")

#forecast from SLT + Random walk
f4 <- fit_stl_2 %>% forecast(method="naive") %>%
  autoplot() + ylab("S01: Var02")

#forecast from Holt-Winters
hw1 <- HoltWinters(s01_ts[,1])

#forecast from Holt-Winters
hw2 <- HoltWinters(s01_ts[,2])

f5 <- hw1 %>% forecast() %>%
  autoplot() + ylab("S01: Var01")

f6 <- hw2 %>% forecast() %>%
  autoplot() + ylab("S01: Var02")

(f1 + f2) / (f3 + f4) / (f5 + f6)
```

#### Forecasting S02: Var02 & Var03 with decomposition

```{r fig.height=10, fig.width=15}
#STL using default values
fit_stl_1 <- stl(s02_ts[,1], s.window = "periodic")

#STL using default values
fit_stl_2 <- stl(s02_ts[,2], s.window = "periodic")

#forecast of seasonaly adjusted data
f1 <- fit_stl_1 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S02: Var02")

#forecast of seasonaly adjusted data
f2 <- fit_stl_2 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S02: Var03")

#forecast from SLT + Random walk
f3 <- fit_stl_1 %>% forecast(method="naive") %>%
  autoplot() + ylab("S02: Var02")

#forecast from SLT + Random walk
f4 <- fit_stl_2 %>% forecast(method="naive") %>%
  autoplot() + ylab("S02: Var03")

#forecast from Holt-Winters
hw1 <- HoltWinters(s02_ts[,1])

#forecast from Holt-Winters
hw2 <- HoltWinters(s02_ts[,2])

f5 <- hw1 %>% forecast() %>%
  autoplot() + ylab("S02: Var02")

f6 <- hw2 %>% forecast() %>%
  autoplot() + ylab("S02: Var03")

(f1 + f2) / (f3 + f4) / (f5 + f6)
```

#### Forecasting S03: Var05 & Var07 with decomposition

```{r fig.height=10, fig.width=15}
#STL using default values
fit_stl_1 <- stl(s03_ts[,1], s.window = "periodic")

#STL using default values
fit_stl_2 <- stl(s03_ts[,2], s.window = "periodic")

#forecast of seasonaly adjusted data
f1 <- fit_stl_1 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S03: Var05")

#forecast of seasonaly adjusted data
f2 <- fit_stl_2 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S03: Var07")

#forecast from SLT + Random walk
f3 <- fit_stl_1 %>% forecast(method="naive") %>%
  autoplot() + ylab("S03: Var05")

#forecast from SLT + Random walk
f4 <- fit_stl_2 %>% forecast(method="naive") %>%
  autoplot() + ylab("S03: Var07")

#forecast from Holt-Winters
hw1 <- HoltWinters(s03_ts[,1])

#forecast from Holt-Winters
hw2 <- HoltWinters(s03_ts[,2])

f5 <- hw1 %>% forecast() %>%
  autoplot() + ylab("S03: Var05")

f6 <- hw2 %>% forecast() %>%
  autoplot() + ylab("S03: Var07")

(f1 + f2) / (f3 + f4) / (f5 + f6)
```

#### Forecasting S04: Var01 & Var02 with decomposition

```{r fig.height=10, fig.width=15}
#STL using default values
fit_stl_1 <- stl(s04_ts[,1], s.window = "periodic")

#STL using default values
fit_stl_2 <- stl(s04_ts[,2], s.window = "periodic")

#forecast of seasonaly adjusted data
f1 <- fit_stl_1 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S04: Var01")

#forecast of seasonaly adjusted data
f2 <- fit_stl_2 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S04: Var02")

#forecast from SLT + Random walk
f3 <- fit_stl_1 %>% forecast(method="naive") %>%
  autoplot() + ylab("S04: Var01")

#forecast from SLT + Random walk
f4 <- fit_stl_2 %>% forecast(method="naive") %>%
  autoplot() + ylab("S04: Var02")

#forecast from Holt-Winters
hw1 <- HoltWinters(s04_ts[,1])

#forecast from Holt-Winters
hw2 <- HoltWinters(s04_ts[,2])

f5 <- hw1 %>% forecast() %>%
  autoplot() + ylab("S04: Var01")

f6 <- hw2 %>% forecast() %>%
  autoplot() + ylab("S04: Var02")

(f1 + f2) / (f3 + f4) / (f5 + f6)
```

#### Forecasting S05: Var02 & Var03 with decomposition

```{r fig.height=10, fig.width=15}
#STL using default values
fit_stl_1 <- stl(s05_ts[,1], s.window = "periodic")

#STL using default values
fit_stl_2 <- stl(s05_ts[,2], s.window = "periodic")

#forecast of seasonaly adjusted data
f1 <- fit_stl_1 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S05: Var02")

#forecast of seasonaly adjusted data
f2 <- fit_stl_2 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S05: Var03")

#forecast from SLT + Random walk
f3 <- fit_stl_1 %>% forecast(method="naive") %>%
  autoplot() + ylab("S05: Var02")

#forecast from SLT + Random walk
f4 <- fit_stl_2 %>% forecast(method="naive") %>%
  autoplot() + ylab("S05: Var03")

#forecast from Holt-Winters
hw1 <- HoltWinters(s05_ts[,1])

#forecast from Holt-Winters
hw2 <- HoltWinters(s05_ts[,2])

f5 <- hw1 %>% forecast() %>%
  autoplot() + ylab("S05: Var02")

f6 <- hw2 %>% forecast() %>%
  autoplot() + ylab("S05: Var03")

(f1 + f2) / (f3 + f4) / (f5 + f6)
```

#### Forecasting S06: Var05 & Var07 with decomposition

```{r fig.height=10, fig.width=15}
#STL using default values
fit_stl_1 <- stl(s06_ts[,1], s.window = "periodic")

#STL using default values
fit_stl_2 <- stl(s06_ts[,2], s.window = "periodic")

#forecast of seasonaly adjusted data
f1 <- fit_stl_1 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S06: Var05")

#forecast of seasonaly adjusted data
f2 <- fit_stl_2 %>% seasadj() %>% naive()%>%
  autoplot() + ylab("S06: Var07")

#forecast from SLT + Random walk
f3 <- fit_stl_1 %>% forecast(method="naive") %>%
  autoplot() + ylab("S06: Var05")

#forecast from SLT + Random walk
f4 <- fit_stl_2 %>% forecast(method="naive") %>%
  autoplot() + ylab("S06: Var07")

#forecast from Holt-Winters
hw1 <- HoltWinters(s06_ts[,1])

#forecast from Holt-Winters
hw2 <- HoltWinters(s06_ts[,2])

f5 <- hw1 %>% forecast() %>%
  autoplot() + ylab("S06: Var05")

f6 <- hw2 %>% forecast() %>%
  autoplot() + ylab("S06: Var07")

(f1 + f2) / (f3 + f4) / (f5 + f6)
```

### Forecasting with Simple Exponential Smoothing (SES)

### Forecasting 

### Forecasting S01: Var01 & Var02 with ses

```{r ses_s01, fig.height=10, fig.width=15}
# simple exponential smoothing since there is no clear trend or seasonal pattern
# Fit ses models to Var01 & Var02
fit_s01_ses_1 <- ses(s01_ts[,1])
fit_s01_ses_2 <- ses(s01_ts[,2])
# check residuals
checkresiduals(fit_s01_ses_1)
checkresiduals(fit_s01_ses_2)
# forecast
fc_s01_ses_1 <- forecast(fit_s01_ses_1)
fc_s01_ses_2 <- forecast(fit_s01_ses_2)
# plot forecasts
fses_S01_1 <- autoplot(fc_s01_ses_1) + ylab("S01: Var01") +
  autolayer(fitted(fc_s01_ses_1))
fses_S01_2 <- autoplot(fc_s01_ses_2) + ylab("S01: Var02") +
  autolayer(fitted(fc_s01_ses_2))
(fses_S01_1 + fses_S01_2)
```


#### Forecasting S02: Var02 & Var03 with ses

```{r ses_s02, fig.height=10, fig.width=15}
# simple exponential smoothing since there is no clear trend or seasonal pattern
# Fit ses models to Var02 & Var03
fit_s02_ses_2 <- ses(s02_ts[,1])
fit_s02_ses_3 <- ses(s02_ts[,2])
# check residuals
checkresiduals(fit_s02_ses_2)
checkresiduals(fit_s02_ses_3)
# forecast
fc_s02_ses_2 <- forecast(fit_s02_ses_2)
fc_s02_ses_3 <- forecast(fit_s02_ses_3)
# plot forecasts
fses_S02_2 <- autoplot(fc_s02_ses_2) + ylab("S02: Var02") +
  autolayer(fitted(fc_s02_ses_2))
fses_S02_3 <- autoplot(fc_s02_ses_3) + ylab("S02: Var03") +
  autolayer(fitted(fc_s02_ses_3))
(fses_S02_2 + fses_S02_3)
```

#### Forecasting S03: Var05 & Var07 with ses

```{r ses_s03, fig.height=10, fig.width=15}
# simple exponential smoothing since there is no clear trend or seasonal pattern
# Fit ses models to Var05 & Var07
fit_s03_ses_5 <- ses(s03_ts[,1])
fit_s03_ses_7 <- ses(s03_ts[,2])
# check residuals
checkresiduals(fit_s03_ses_5)
checkresiduals(fit_s03_ses_7)
# forecast
fc_s03_ses_5 <- forecast(fit_s03_ses_5)
fc_s03_ses_7 <- forecast(fit_s03_ses_7)
# plot forecasts
fses_S03_5 <- autoplot(fc_s03_ses_5) + ylab("S03: Var05") +
  autolayer(fitted(fc_s03_ses_5))
fses_S03_7 <- autoplot(fc_s03_ses_7) + ylab("S03: Var07") +
  autolayer(fitted(fc_s03_ses_7))
(fses_S03_5 + fses_S03_7)
```

#### Forecasting S04: Var01 & Var02 with ses

```{r ses_s04, fig.height=10, fig.width=15}
# simple exponential smoothing since there is no clear trend or seasonal pattern
# Fit ses models to Var01 & Var02
fit_s04_ses_1 <- ses(s04_ts[,1])
fit_s04_ses_2 <- ses(s04_ts[,2])
# check residuals
checkresiduals(fit_s04_ses_1)
checkresiduals(fit_s04_ses_2)
# forecast next 140 periods
fc_s04_ses_1 <- forecast(fit_s04_ses_1)
fc_s04_ses_2 <- forecast(fit_s04_ses_2)
# plot forecasts
fses_s04_1 <- autoplot(fc_s04_ses_1) + ylab("s04: Var01") +
  autolayer(fitted(fc_s04_ses_1))
fses_s04_2 <- autoplot(fc_s04_ses_2) + ylab("s04: Var02") +
  autolayer(fitted(fc_s04_ses_2))
(fses_s04_1 + fses_s04_2)
```


#### Forecasting S05: Var02 & Var03 with ses

```{r ses_s05, fig.height=10, fig.width=15}
# simple exponential smoothing since there is no clear trend or seasonal pattern
# Fit ses models to Var02 & Var03
fit_s05_ses_2 <- ses(s05_ts[,1])
fit_s05_ses_3 <- ses(s05_ts[,2])
# check residuals
checkresiduals(fit_s05_ses_2)
checkresiduals(fit_s05_ses_3)
# forecast
fc_s05_ses_2 <- forecast(fit_s05_ses_2)
fc_s05_ses_3 <- forecast(fit_s05_ses_3)
# plot forecasts
fses_s05_2 <- autoplot(fc_s05_ses_2) + ylab("s05: Var02") +
  autolayer(fitted(fc_s05_ses_2))
fses_s05_3 <- autoplot(fc_s05_ses_3) + ylab("s05: Var03") +
  autolayer(fitted(fc_s05_ses_3))
(fses_s05_2 + fses_s05_3)
```


#### Forecasting S06: Var05 & Var07 with ses

```{r ses_s06, fig.height=10, fig.width=15}
# simple exponential smoothing since there is no clear trend or seasonal pattern
# Fit ses models to Var05 & Var07
fit_s06_ses_5 <- ses(s06_ts[,1])
fit_s06_ses_7 <- ses(s06_ts[,2])
# check residuals
checkresiduals(fit_s06_ses_5)
checkresiduals(fit_s06_ses_7)
# forecast
fc_s06_ses_5 <- forecast(fit_s06_ses_5)
fc_s06_ses_7 <- forecast(fit_s06_ses_7)
# plot forecasts
fses_s06_5 <- autoplot(fc_s06_ses_5) + ylab("s06: Var05") +
  autolayer(fitted(fc_s06_ses_5))
fses_s06_7 <- autoplot(fc_s06_ses_7) + ylab("s06: Var07") +
  autolayer(fitted(fc_s06_ses_7))
(fses_s06_5 + fses_s06_7)
```


### Forecasting using ETS models

#### Forecasting S01: Var01 & Var02 with ETS

```{r ets_s01, fig.height=10, fig.width=15}
# Fit ETS models to Var01 & Var02
fit_s01_ETS_1 <- ets(s01_ts[,1])
fit_s01_ETS_2 <- ets(s01_ts[,2])
# check residuals
checkresiduals(fit_s01_ETS_1)
checkresiduals(fit_s01_ETS_2)
# forecast
fc_s01_ETS_1 <- forecast(fit_s01_ETS_1)
fc_s01_ETS_2 <- forecast(fit_s01_ETS_2)
# plot forecasts
fets_S01_1 <- autoplot(fc_s01_ETS_1) + ylab("S01: Var01") +
  autolayer(fitted(fc_s01_ETS_1))
fets_S01_2 <- autoplot(fc_s01_ETS_2) + ylab("S01: Var02") +
  autolayer(fitted(fc_s01_ETS_2))
(fets_S01_1 + fets_S01_2)
```

#### Forecasting S02: Var02 & Var03 with ETS

```{r ets_s02, fig.height=10, fig.width=15}
# Fit ETS models to Var02 & Var03
fit_s02_ETS_2 <- ets(s02_ts[,1])
fit_s02_ETS_3 <- ets(s02_ts[,2])
# check residuals
checkresiduals(fit_s02_ETS_2)
checkresiduals(fit_s02_ETS_3)
# forecast
fc_s02_ETS_2 <- forecast(fit_s02_ETS_2)
fc_s02_ETS_3 <- forecast(fit_s02_ETS_3)
# plot forecasts
fets_S02_2 <- autoplot(fc_s02_ETS_2) + ylab("S02: Var02") +
  autolayer(fitted(fc_s02_ETS_2))
fets_S02_3 <- autoplot(fc_s02_ETS_3) + ylab("S02: Var03") +
  autolayer(fitted(fc_s02_ETS_3))
(fets_S02_2 + fets_S02_3)
```


#### Forecasting S03: Var05 & Var07 with ETS

```{r ets_s03, fig.height=10, fig.width=15}
# Fit ETS models to Var05 & Var07
fit_s03_ETS_5 <- ets(s03_ts[,1])
fit_s03_ETS_7 <- ets(s03_ts[,2])
# check residuals
checkresiduals(fit_s03_ETS_5)
checkresiduals(fit_s03_ETS_7)
# forecast
fc_s03_ETS_5 <- forecast(fit_s03_ETS_5)
fc_s03_ETS_7 <- forecast(fit_s03_ETS_7)
# plot forecasts
fets_S03_5 <- autoplot(fc_s03_ETS_5) + ylab("S03: Var05") +
  autolayer(fitted(fc_s03_ETS_5))
fets_S03_7 <- autoplot(fc_s03_ETS_7) + ylab("S03: Var07") +
  autolayer(fitted(fc_s03_ETS_7))
(fets_S03_5 + fets_S03_7)
```


#### Forecasting S04: Var01 & Var02 with ETS

```{r ets_s04, fig.height=10, fig.width=15}
# Fit ETS models to Var01 & Var02
fit_s04_ETS_1 <- ets(s04_ts[,1])
fit_s04_ETS_2 <- ets(s04_ts[,2])
# check residuals
checkresiduals(fit_s04_ETS_1)
checkresiduals(fit_s04_ETS_2)
# forecast next 140 periods
fc_s04_ETS_1 <- forecast(fit_s04_ETS_1)
fc_s04_ETS_2 <- forecast(fit_s04_ETS_2)
# plot forecasts
fets_s04_1 <- autoplot(fc_s04_ETS_1) + ylab("s04: Var01") +
  autolayer(fitted(fc_s04_ETS_1))
fets_s04_2 <- autoplot(fc_s04_ETS_2) + ylab("s04: Var02") +
  autolayer(fitted(fc_s04_ETS_2))
(fets_s04_1 + fets_s04_2)
```


#### Forecasting S05: Var02 & Var03 with ETS

```{r ets_s05, fig.height=10, fig.width=15}
# Fit ETS models to Var02 & Var03
fit_s05_ETS_2 <- ets(s05_ts[,1])
fit_s05_ETS_3 <- ets(s05_ts[,2])
# check residuals
checkresiduals(fit_s05_ETS_2)
checkresiduals(fit_s05_ETS_3)
# forecast
fc_s05_ETS_2 <- forecast(fit_s05_ETS_2)
fc_s05_ETS_3 <- forecast(fit_s05_ETS_3)
# plot forecasts
fets_s05_2 <- autoplot(fc_s05_ETS_2) + ylab("s05: Var02") +
  autolayer(fitted(fc_s05_ETS_2))
fets_s05_3 <- autoplot(fc_s05_ETS_3) + ylab("s05: Var03") +
  autolayer(fitted(fc_s05_ETS_3))
(fets_s05_2 + fets_s05_3)
```


#### Forecasting S06: Var05 & Var07 with ETS

```{r ets_s06, fig.height=10, fig.width=15}
# Fit ETS models to Var05 & Var07
fit_s06_ETS_5 <- ets(s06_ts[,1])
fit_s06_ETS_7 <- ets(s06_ts[,2])
# check residuals
checkresiduals(fit_s06_ETS_5)
checkresiduals(fit_s06_ETS_7)
# forecast
fc_s06_ETS_5 <- forecast(fit_s06_ETS_5)
fc_s06_ETS_7 <- forecast(fit_s06_ETS_7)
# plot forecasts
fets_s06_5 <- autoplot(fc_s06_ETS_5) + ylab("s06: Var05") +
  autolayer(fitted(fc_s06_ETS_5))
fets_s06_7 <- autoplot(fc_s06_ETS_7) + ylab("s06: Var07") +
  autolayer(fitted(fc_s06_ETS_7))
(fets_s06_5 + fets_s06_7)
```


<br>
<br>

### Forecasting using ARIMA models

#### Forecasting S01: Var01 & Var02 with ARIMA

```{r arima_s01, fig.height=10, fig.width=15}
# Fit ARIMA models to Var01 & Var02
fit_s01_ARIMA_1 <- auto.arima(s01_ts[,1], stepwise = TRUE)
fit_s01_ARIMA_2 <- auto.arima(s01_ts[,2], stepwise = TRUE)
# check residuals
checkresiduals(fit_s01_ARIMA_1)
checkresiduals(fit_s01_ARIMA_2)
# forecast
fc_s01_ARIMA_1 <- forecast(fit_s01_ARIMA_1)
fc_s01_ARIMA_2 <- forecast(fit_s01_ARIMA_2)
# plot forecasts
fa_S01_1 <- autoplot(fc_s01_ARIMA_1) + ylab("S01: Var01") +
  autolayer(fitted(fc_s01_ARIMA_1))
fa_S01_2 <- autoplot(fc_s01_ARIMA_2) + ylab("S01: Var02") +
  autolayer(fitted(fc_s01_ARIMA_2))
(fa_S01_1 + fa_S01_2)
```


#### Forecasting S02: Var02 & Var03 with ARIMA

```{r arima_s02, fig.height=10, fig.width=15}
# Fit ARIMA models to Var02 & Var03
fit_s02_ARIMA_2 <- auto.arima(s02_ts[,1], stepwise = TRUE)
fit_s02_ARIMA_3 <- auto.arima(s02_ts[,2], stepwise = TRUE)
# check residuals
checkresiduals(fit_s02_ARIMA_2)
checkresiduals(fit_s02_ARIMA_3)
# forecast
fc_s02_ARIMA_2 <- forecast(fit_s02_ARIMA_2)
fc_s02_ARIMA_3 <- forecast(fit_s02_ARIMA_3)
# plot forecasts
fa_S02_2 <- autoplot(fc_s02_ARIMA_2) + ylab("S02: Var02") + ylab("S01: Var01") +
  autolayer(fitted(fc_s02_ARIMA_2))
fa_S02_3 <- autoplot(fc_s02_ARIMA_3) + ylab("S02: Var03") + ylab("S01: Var01") +
  autolayer(fitted(fc_s02_ARIMA_3))
(fa_S02_2 + fa_S02_3)
```

#### Forecasting S03: Var05 & Var07 with ARIMA

```{r arima_s03, fig.height=10, fig.width=15}
# Fit ARIMA models to Var05 & Var07
fit_s03_ARIMA_5 <- auto.arima(s03_ts[,1], stepwise = TRUE)
fit_s03_ARIMA_7 <- auto.arima(s03_ts[,2], stepwise = TRUE)
# check residuals
checkresiduals(fit_s03_ARIMA_5)
checkresiduals(fit_s03_ARIMA_7)
# forecast
fc_s03_ARIMA_5 <- forecast(fit_s03_ARIMA_5)
fc_s03_ARIMA_7 <- forecast(fit_s03_ARIMA_7)
# plot forecasts
fa_S03_5 <- autoplot(fc_s03_ARIMA_5) + ylab("S03: Var05") + ylab("S01: Var01") +
  autolayer(fitted(fc_s03_ARIMA_5))
fa_S03_7 <- autoplot(fc_s03_ARIMA_7) + ylab("S03: Var07") + ylab("S01: Var01") +
  autolayer(fitted(fc_s03_ARIMA_7))
(fa_S03_5 + fa_S03_7)
```


#### Forecasting S04: Var01 & Var02 with ARIMA

```{r arima_s04, fig.height=10, fig.width=15}
# Fit ARIMA models to Var01 & Var02
fit_s04_ARIMA_1 <- auto.arima(s04_ts[,1], stepwise = TRUE)
fit_s04_ARIMA_2 <- auto.arima(s04_ts[,2], stepwise = TRUE)
# check residuals
checkresiduals(fit_s04_ARIMA_1)
checkresiduals(fit_s04_ARIMA_2)
# forecast
fc_s04_ARIMA_1 <- forecast(fit_s04_ARIMA_1)
fc_s04_ARIMA_2 <- forecast(fit_s04_ARIMA_2)
# plot forecasts
fa_S04_1 <- autoplot(fc_s04_ARIMA_1) + ylab("S04: Var01") + ylab("S01: Var01") +
  autolayer(fitted(fc_s04_ARIMA_1))
fa_S04_2 <- autoplot(fc_s04_ARIMA_2) + ylab("S04: Var02") + ylab("S01: Var01") +
  autolayer(fitted(fc_s04_ARIMA_2))
(fa_S04_1 + fa_S04_2)
```

#### Forecasting S05: Var02 & Var03 with ARIMA

```{r arima_s05, fig.height=10, fig.width=15}
# Fit ARIMA models to Var02 & Var03
fit_s05_ARIMA_2 <- auto.arima(s05_ts[,1], stepwise = TRUE)
fit_s05_ARIMA_3 <- auto.arima(s05_ts[,2], stepwise = TRUE)
# check residuals
checkresiduals(fit_s05_ARIMA_2)
checkresiduals(fit_s05_ARIMA_3)
# forecast
fc_s05_ARIMA_2 <- forecast(fit_s05_ARIMA_2)
fc_s05_ARIMA_3 <- forecast(fit_s05_ARIMA_3)
# plot forecasts
fa_S05_2 <- autoplot(fc_s05_ARIMA_2) + ylab("S05: Var02") + ylab("S01: Var01") +
  autolayer(fitted(fc_s05_ARIMA_2))
fa_S05_3 <- autoplot(fc_s05_ARIMA_3) + ylab("S05: Var03") + ylab("S01: Var01") +
  autolayer(fitted(fc_s05_ARIMA_3))
(fa_S05_2 + fa_S05_3)
```


#### Forecasting S06: Var05 & Var07 with ARIMA

```{r arima_s06, fig.height=10, fig.width=15}
# Fit ARIMA models to Var05 & Var07
fit_s06_ARIMA_5 <- auto.arima(s06_ts[,1], stepwise = TRUE)
fit_s06_ARIMA_7 <- auto.arima(s06_ts[,2], stepwise = TRUE)
# check residuals
checkresiduals(fit_s06_ARIMA_5)
checkresiduals(fit_s06_ARIMA_7)
# forecast
fc_s06_ARIMA_5 <- forecast(fit_s06_ARIMA_5)
fc_s06_ARIMA_7 <- forecast(fit_s06_ARIMA_7)
# plot forecasts
fa_S06_5 <- autoplot(fc_s06_ARIMA_5) + ylab("S06: Var05") + ylab("S01: Var01") +
  autolayer(fitted(fc_s06_ARIMA_5))
fa_S06_7 <- autoplot(fc_s06_ARIMA_7) + ylab("S06: Var07") + ylab("S01: Var01") +
  autolayer(fitted(fc_s06_ARIMA_7))
(fa_S06_5 + fa_S06_7)
```


## Model Selection


## Final Forecasting



