---
title: "Data624 - Project 1"
author: "Esteban Aramayo, Coffy Andrews-Guo, LeTicia Cancel, Joseph Connolly, Ian Costello"
date: '2022-06-21'
output: 
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE,
                      message=FALSE,
                      collapse = FALSE,
                      comment = "#>" )
```

```{r echo=FALSE}
# install from CRAN
#install.packages("officedown")

# or GitHub
#remotes::install_github("davidgohel/officedown")
```

Required Libraries
```{r warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(fpp2)
library(caret)
library(RANN)
library(VIM)
library(ggpubr)
library(gridExtra)
library(forecast)
library(writexl)
```

```{r}
df <- read_excel("data.xls")
head(df)
```

## Data Cleaning & Imputation

```{r}
# Factoring category to get a count of the elements within dataset
df$category <- as.factor(df$category)


summary(df)
writeLines("\n\n")
str(df)
```

```{r}
df <- mice(df, method = "cart")
df <- complete(df)
```

paste0(sum(is.na(df))," values missing after imputation")
```

```{r}
# For forecasting later on
#md

s01 <- df %>% filter(category == "S01")
s02 <- df %>% filter(category == "S02")
s03 <- df %>% filter(category == "S03")
s04 <- df %>% filter(category == "S04")
s05 <- df %>% filter(category == "S05")
s06 <- df %>% filter(category == "S06")
```

```{r}
s01_ts <- ts(s01[,c("Var01","Var02")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s02_ts <- ts(s02[,c("Var02","Var03")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s03_ts <- ts(s03[,c("Var05","Var07")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s04_ts <- ts(s04[,c("Var01","Var02")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s05_ts <- ts(s05[,c("Var02","Var03")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
s06_ts <- ts(s06[,c("Var05","Var07")], frequency = 12, start = c(2011, 5), end = c(2018, 5))
```

```{r arima_s01, fig.height=10, fig.width=15}
# Fit ARIMA models to Var01 & Var02
fit_s01_ARIMA_1 <- auto.arima(s01_ts[,1])
fit_s01_ARIMA_2 <- auto.arima(s01_ts[,2])
# check residuals
checkresiduals(fit_s01_ARIMA_1)
checkresiduals(fit_s01_ARIMA_2)
# forecast
fc_s01_ARIMA_1 <- forecast(fit_s01_ARIMA_1)
fc_s01_ARIMA_2 <- forecast(fit_s01_ARIMA_2)
# plot forecasts
fa_S01_1 <- autoplot(fc_s01_ARIMA_1) + ylab("S01: Var01") +
  autolayer(fitted(fc_s01_ARIMA_1))
fa_S01_2 <- autoplot(fc_s01_ARIMA_2) + ylab("S01: Var02") +
  autolayer(fitted(fc_s01_ARIMA_2))
(fa_S01_1 + fa_S01_2)
```

